\sysfull

### 2017-08-10 paper/compiler.tex

{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block. Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we simply use the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost. Moreover, we assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy. This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement. \sys's analysis tries to ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget (which is part of the boundary placement process described later).

{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of variable that {\em may} be {\em alive} across the edge (i.e., variables whose value may be consumed later on). Variables that live across an edge must be protected by \sys's memory virtualization mechanism and each protected, inter-task variable increases the run time overhead for a boundary placement. \sys's tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.

{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
annotated with a static loop bound. For loops that do not have a static bound, \sys conservatively assumes the maximum possible loop bound, based on the type of the loop induction variable.

@@

\sys uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score. The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of variables live variables at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero and one, and are hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
$N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across applications. We found that varying weights varies runtime by only around 5\% in most cases and in our experimental evaluation, we used $w_{0} = w_{1} = 0.5$.

>>>>

{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block. Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we simply use the number of intermediate representation (IR) instructions in a block as a proxy for energy because more instructions mean more energy consumption. Moreover, we assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy. This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement. \sys's analysis tries to ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget (which is part of the boundary placement process described later).

{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of variable that {\em may} be {\em alive} across the edge (i.e., variables whose value may be consumed later on). Variables that live across an edge must be protected by \sys's memory virtualization mechanism when boundary is placed at the edge and the resulting inter-task variable increases the run time overhead. \sys's tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.

{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
annotated with a static loop bound. For loops that do not have a static bound, \sys conservatively assumes the loop will be too large to fit within a single task.

@@

\sys uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score. The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path. A boundary candidate's score is computed as
%
\begin{equation}
\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
\end{equation}
%
where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of protected variables that will be newly introduced by placing boundary at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero and one, and are hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
$N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across applications. We found that varying weights varies runtime by only around 5\% in most cases and in our experimental evaluation, we used $w_{0} = w_{1} = 0.5$.

### 2017-08-10 paper/evaluation.tex

We also used an alternative setup where the MSP-TS430RGZ48C launchpad was for continuous power experiments. We used internal timer for measuring run time. For harvested energy, we used the ThingMagic Astra-EX RFID reader with power levels in the [17.25,30]\,dBm range. We used EDB~\cite{edb} for execution time measurement. Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.\todo{Report on LLVM version}{Brandon}

@@

Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile DFT application since it was using built-ins for C math functions (i.e. \texttt{rand()}, \texttt{sinf()}, \texttt{cosf()}) that were not provided by the Clang MSP430 back-end.}. The result are given in Table~\ref{table:compile_time}. Most of the compile time was due to pointer aliasing.\todo{Comment on: whether these times are good or bad; why there is such a difference in time between all applications; how many runs you performed; what PC/OS you used to run the code; how the result was measured; why results are rounded (there are no decimals in your result); what is "Clang MSP430 back-end"}{Kiwan}

\begin{table}[t]
	\centering
	\footnotesize
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		Bitcount & Cold chain & Cuckoo & RSA & AR & Sort \\
		\hline\hline
		3 & 2 & 6 & 86 & 34 & > 1 \\
		\hline
	\end{tabular}
	\caption{\sys compile time per application (seconds).\todo{for Sort:">" or "<"?}{Kiwan}}
\label{table:compile_time}
\end{table}

>>>>

We also used an alternative setup where the MSP-TS430RGZ48C launchpad was for continuous power experiments. We used internal timer for measuring run time. For harvested energy, we used the ThingMagic Astra-EX RFID reader with power levels of 30\,dBm. We used EDB~\cite{edb} for execution time measurement. Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler. Both for Clang and LLVM we used version 3.8.0.

@@

Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile DFT application since the LLVM inserts mathematical built-in functions (i.e. \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}. We compiled the apps on a 4-core virtual machine with 1GB memory that has Arch Linux, running on Intel Core i5-6200U (2.30GHz) processor. The result are given in Table~\ref{table:compile_time}. We simply used {\em date} command provided by linux to measure the rough time. Most of the compile time was due to pointer aliasing, when calculating possible live range of each variables. Apps such as RSA whose code size is large and uses a lot of variables needed longer compilation time due to larger search space for pointer aliasing.

\begin{table}[t]
	\centering
	\footnotesize
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		Bitcount & Cold chain & Cuckoo & RSA & AR & Sort \\
		\hline\hline
		3 & 2 & 6 & 86 & 34 & < 1 \\
		\hline
	\end{tabular}
	\caption{\sys compile time per application (seconds).}
\label{table:compile_time}
\end{table}
